### 2018-11-30 Revision/Paper/conc.tex
..

%%Brief summary
%-Mathematical connections between Sobol, DGSMs, and activity scores exist and could
%be exploited to perform GSA with reduced effort.
%-Iterative implementation yields advantage
%-Consistency between the two approaches with DGSM as well as among themselves
%encourages the use of gradient-free.
%-Grad-free was found to be suitable for a high-dimensional problem.
%-Based on the results presented, the following parameter were found to be 
%significant.
%
%discussion
%- Methodology is agnostic to the choice of mechanism and could be applied to other
%systems.
%- However, constraints exist in terms of differentiability of the QoI. Not all
%input/output relationships exhibit an active subspace. However, it is important to
%check.
%- Activity scores could be used to identify key reactions and thereby construct
%reduced order mechanisms for reducing computational effort and investigating the
%inadequacy associated with rate laws.

In this work, we focused on the uncertainty associated with the
rate-controlling parameters in the H$_2$/O$_2$ reaction mechanism and its
impact on ignition delay predictions. The mechanism involves 19 different
reactions and in each case, the reaction rate depends upon the choice of a
pre-exponent and an activation energy. Hence, in theory, the evolution of the
chemical system depends upon 38 inputs. However, we considered 
epistemic uncertainty in all pre-exponents and activation energies with non-zero
nominal values i.e. a total of 33 parameters instead of 38.  
%Conventional means for uncertainty quantification such as those involving
%surrogate model construction as well as sensitivity analysis would thus be
%computationally challenging.
To facilitate efficient uncertainty analysis, we focused our efforts on
reducing the dimensionality of the problem by identifying important directions
in the parameter space such that the model output 
predominantly varies along these directions. These important directions
constitute the active subspace. Additionally, we demonstrated that the activity scores,
computed using the components of the dominant eigenvectors provide an efficient
means for approximating derivative based global sensitivity measures (DGSMs).
Furthermore, we established generalized mathematical linkages between the
different global sensitivity measures: activity scores, DGSMs, and total Sobol'
index which could be exploited to reduce computational effort associated with
global sensitivity analysis.

Active subspace computation requires repeated evaluations of the gradient of
the QoI i.e. the ignition delay. For this purpose, we explored two approaches,
namely, perturbation-based (e.g. finite difference - used in this work, adjoint-based
methods, etc.) and regression-based. Both approaches were shown to yield consistent
results for the 19-dimensional
problem wherein only the pre-exponents were considered to be uncertain.
Additionally, the activity scores were also shown to be consistent with the
screening metric estimates based on DGSMs in~\cite{Vohra:2018}. An
iterative procedure was adopted to enhance the computational efficiency.

The active subspace was further computed for a 33-dimensional problem
wherein all pre-exponents and activation energies with non-zero nominal
estimates were considered uncertain. Once again, the active subspace
was found to be 1-dimensional. However, unlike the 19-dimensional case,
gradient estimation using the regression-based approach resulted in an
under-estimation of the variability in the ignition delay and also failed
to capture the sensitivity towards one of the most important parameters, $E_{a,15}$.
We attribute
this discrepancy to two potential sources of numerical errors: (1)
Regression-based approximation of the gradient of the model output, and (2)
Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
error incurred by the linear regression fit in this case. 
Global sensitivity estimates using the perturbation-based
approximation of the gradient indicated that the variability in the ignition
delay is  largely dependent on $A_1$, $A_9$,
$E_{a,1}$ and $E_{a,15}$, although contributions from $A_{15}$ and $A_{17}$
were also found to be significant. 
Hence, while the
regression-based approach 
reduces computational effort, the present investigations clearly reveal
potential pitfalls associated with this approach.

%Computational gains using the proposed framework in this work are essentially
%realized by using the 1-dimensional surrogate in the active subspace that
%essentially casts a mutlivariate input in the full space into a univariate
%input in the reduced space. The surrogate can be used to accelerate forward
%propagation of the uncertainty as well as parameter estimation in a Bayesian
%setting. Hence, we focused our efforts to verify the accuracy of the
%1-dimensional surrogates, obtained using both strategies.  The 1-dimensional
%surrogates were assessed for accuracy by first evaluating the relative $L^2$-norm
%of the error between its predictions and model evaluations at $10^4$
%independent MC samples in the 33-dimensional input domain. The relative error
%computed using predictions from the surrogate resulting from the regression-based
%approach were found to be an order of magnitude larger than that estimated
%using surrogate resulting from the perturbation-based approximation of the gradient.

%Additionally, the surrogates were assessed by comparing probability
%distributions for the ignition delay,
%constructed using surrogate predictions and model evaluations at the same set
%of 10$^4$ random samples. The PDFs based on model evaluations and the
%1-dimensional surrogate from the perturbation-based approach were found to be
%almost identical. However, the PDF based on 1-dimensional surrogate from the
%regression-based approach, although captured the modal estimate with reasonable
%accuracy, was observed to underestimate the uncertainty in the ignition delay.
%Mean estimates of the ignition delay using model evaluations and predictions
%using the two surrogates at 10$^4$ samples were found to be in agreement.
%However, the standard deviation, computed using predictions from
%the surrogate resulting from the regression-based approach was found to be accurate
%only upto the first significant digit. Moreover, it was observed that the
%activity scores, computed using the regression-based strategy failed to capture
%the sensitivity towards one of the most important parameters, $E_{a,15}$.  
%
%More specifically, using
%the regression-based approach, 
%\begin{itemize} 
%\item The desired statistics (e.g. mean and variance) of the
%QoI might not be approximated to a desired level of accuracy.  
%\item The
%approach might fail to capture the sensitivity  of the QoI towards all inputs.
%\end{itemize} 
Therefore, in general, perturbation-based approaches should be preferred for 
active subspace computation.
The regression-based approach can be explored in situations involving intensive
simulations where the perturbation-based methods are intractable.
%
%and the goal is obtaining rough estimates of the statistics of the QoI as
%opposed to a more detailed analysis such as those involving parametric
%sensitivity.  
%
We also mention that alternate regression based approaches such as ones based
on computing a global quadratic model have been proposed and used in the
literature; see e.g.,~\cite{Constantine:2017a}.  The applicability of such an
approach in the context of high-dimensional chemical reaction networks is
subject to future work.

The computational framework presented in this work is agnostic to the choice of
the chemical system and can be easily adapted for other systems as long
as the quantity of interest is continuously differentiable in the considered
domain of the inputs.  We have demonstrated that the active subspace could be
exploited for efficient forward propagation of the uncertainty from inputs to
the output. The resulting activity scores and the low-dimensional surrogate
could further guide optimal allocation of computational resources for
calibration of the important rate-controlling parameters in a Bayesian setting.
Additionally, dimension reduction using active subspaces could assist in
developing robust formulations for predicting discrepancy between simulations
and measurements due to epistemic uncertainty in the model inputs.

>>>>

In this work, we focused on the uncertainty associated with the
rate-controlling parameters in the H$_2$/O$_2$ reaction mechanism
\rebut{as well as the initial pressure, temperature, and stoichiometry} and its
impact on ignition delay predictions. The mechanism involves 19 different
reactions and in each case, the reaction rate depends upon the choice of a
pre-exponent and an activation energy. Hence, in theory, the evolution of the
chemical system depends upon 38 rate parameters \rebut{and three initial conditions}.
However, we considered 
epistemic uncertainty in all pre-exponents and activation energies with non-zero
nominal values i.e. a total of 33 rate parameters instead of 38 \rebut{in addition to
the three initial conditions}.  
%Conventional means for uncertainty quantification such as those involving
%surrogate model construction as well as sensitivity analysis would thus be
%computationally challenging.
To facilitate efficient uncertainty analysis, we focused our efforts on
reducing the dimensionality of the problem by identifying important directions
in the parameter space such that the model output 
predominantly varies along these directions. These important directions
constitute the active subspace. Additionally, we demonstrated that the activity scores,
computed using the components of the dominant eigenvectors provide an efficient
means for approximating derivative based global sensitivity measures (DGSMs).
Furthermore, we established generalized mathematical linkages between the
different global sensitivity measures: activity scores, DGSMs, and total Sobol'
index which could be exploited to reduce computational effort associated with
global sensitivity analysis.

Active subspace computation requires repeated evaluations of the gradient of
the QoI i.e. the ignition delay. For this purpose, we explored two approaches,
namely, perturbation-based (e.g. finite difference - used in this work, adjoint-based
methods, etc.) and regression-based. Both approaches were shown to yield consistent
results for the 19-dimensional
problem wherein only the pre-exponents were considered to be uncertain.
\rebut{It was observed that the computational effort required to obtain a 
converged active subspace was comparable for the two approaches. However,
the predictive accuracy of the perturbation approach was observed to be
slightly higher. Moreover, a 1-dimensional active subspace was shown to
reasonably approximate the uncertainty in the ignition delay.}
Additionally, the activity scores were also shown to be consistent with the
screening metric estimates based on DGSMs in~\cite{Vohra:2018}. An
iterative procedure was adopted to enhance the computational efficiency.

The active subspace was further computed for a 36-dimensional problem
wherein all pre-exponents and activation energies with non-zero nominal
estimates \rebut{as well as the initial conditions} were considered uncertain. 
\rebut{Once again, consistent results were obtained using the two approaches.
A 1-dimensional active subspace was shown to reasonably
capture the uncertainty in the ignition delay in this case. However, the
computational effort required to compute a converged active subspace
using perturbation was found to be half of the effort required in the case
of regression. Predictive accuracy of the two approaches was found to 
be comparable. Hence, perturbation seems like a preferred approach
for the higher-dimensional problem based on our findings. GSA results indicated
that the variability in the ignition delay is predominantly due to the 
uncertainty in the rate parameters, $A_1$ and $A_9$ with significant
contributions from $A_{15}$, $A_{17}$, and $E_{a,15}$. Additionally, the
ignition delay was found to be sensitive towards $T_0$.}

Therefore, in general, perturbation-based approaches should be preferred for 
active subspace computation especially in situations where efficient gradient
computation techniques (adjoints-based, automatic differentiation, etc.) are
feasible. The regression-based approach can be explored in situations involving intensive
simulations where the perturbation-based methods are intractable.
%
%and the goal is obtaining rough estimates of the statistics of the QoI as
%opposed to a more detailed analysis such as those involving parametric
%sensitivity.  
%
We also mention that alternate regression based approaches such as ones based
on computing a global quadratic model have been proposed and used in the
literature; see e.g.,~\cite{Constantine:2017a}.  The applicability of such an
approach in the context of high-dimensional chemical reaction networks is
subject to future work.

The computational framework presented in this work is agnostic to the choice of
the chemical system and can be easily adapted for other systems as long
as the quantity of interest is continuously differentiable in the considered
domain of the inputs.  We have demonstrated that the active subspace could be
exploited for efficient forward propagation of the uncertainty from inputs to
the output. The resulting activity scores and the low-dimensional surrogate
could further guide optimal allocation of computational resources for
calibration of the important rate-controlling parameters \rebut{and input conditions}
 in a Bayesian setting.
Additionally, dimension reduction using active subspaces could assist in
developing robust formulations for predicting discrepancy between simulations
and measurements due to epistemic uncertainty in the model inputs.

### 2018-11-30 Revision/Paper/intro.tex
..

The application problem considered in this work is the
H$_2$/O$_2$ reaction mechanism from~\cite{Yetter:1991}. This mechanism has gained
a lot of attention as a potential source of clean energy for
locomotive applications~\cite{Das:1996}, and more recently in fuel
 cells~\cite{Loges:2008,Cosnier:2016}. 
The mechanism involves 19
reactions including chain reactions, dissociation/recombination reactions, and
formation and consumption of intermediate species; see Table~\ref{tab:kinetics}. 
For each reaction, the reaction rate is assumed to follow an Arrhenius
correlation with temperature:
%
\be
k_i(T) = A_iT^{n_i}\exp(-E_{a,i}/RT), 
\label{eq:rate}
\ee
%
where $A_i$ is the pre-exponent, $n_i$ is the index of $T$, $E_{a,i}$ is the
activation energy corresponding to the $i^{th}$ reaction, and $R$ is the
universal gas constant.  
%
\begin{table}[htbp]
\renewcommand*{\arraystretch}{0.9}
\begin{center}
\begin{tabular}{llll}
\toprule
Reaction \#     & Reaction &&\\
\bottomrule
$\mathcal{R}_1$ & H + O$_2$          & $\rightleftharpoons$ & O + OH \\
$\mathcal{R}_2$ & O + H$_2$          & $\rightleftharpoons$ & H + OH \\
$\mathcal{R}_3$ & H$_2$ + OH         & $\rightleftharpoons$ & H$_2$O + H \\
$\mathcal{R}_4$ & OH + OH            & $\rightleftharpoons$ & O + H$_2$O \\
$\mathcal{R}_5$ & H$_2$ + M          & $\rightleftharpoons$ & H + H + M \\
$\mathcal{R}_6$ & O + O + M          & $\rightleftharpoons$ & O$_2$ + M \\
$\mathcal{R}_7$ & O + H + M          & $\rightleftharpoons$ & OH + M \\
$\mathcal{R}_8$ & H + OH +M          & $\rightleftharpoons$ & H$_2$O + M \\
$\mathcal{R}_9$ & H + O$_2$ + M      & $\rightleftharpoons$ & HO$_2$ + M \\
$\mathcal{R}_{10}$ & HO$_2$ + H      & $\rightleftharpoons$ & H$_2$ + O$_2$ \\
$\mathcal{R}_{11}$ & HO$_2$ + H      & $\rightleftharpoons$ & OH + OH \\
$\mathcal{R}_{12}$ & HO$_2$ + O      & $\rightleftharpoons$ & O$_2$ + OH \\
$\mathcal{R}_{13}$ & HO$_2$ + OH     & $\rightleftharpoons$ & H$_2$O + O$_2$ \\
$\mathcal{R}_{14}$ & HO$_2$ + HO$_2$ & $\rightleftharpoons$ & H$_2$O$_2$ + O$_2$ \\
$\mathcal{R}_{15}$ & H$_2$O$_2$ + M  & $\rightleftharpoons$ & OH + OH + M \\
$\mathcal{R}_{16}$ & H$_2$O$_2$ + H  & $\rightleftharpoons$ & H$_2$O + OH \\
$\mathcal{R}_{17}$ & H$_2$O$_2$ + H  & $\rightleftharpoons$ & HO$_2$ + H$_2$ \\
$\mathcal{R}_{18}$ & H$_2$O$_2$ + O  & $\rightleftharpoons$ & OH + HO$_2$ \\
$\mathcal{R}_{19}$ & H$_2$O$_2$ + OH & $\rightleftharpoons$ & HO$_2$ + H$_2$O \\
\bottomrule
\end{tabular}
\end{center}
\caption{Reaction mechanism for H$_2$/O$_2$ from~\cite{Yetter:1991}}.
\label{tab:kinetics}
\end{table}
%
The global reaction associated with the H$_2$/O$_2$ mechanism can
be considered as follows:
\be
2\text{H}_2 + \text{O}_2 \rightarrow 2\text{H}_2\text{O}.
\label{eq:global}
\ee 
The equivalence ratio ($\phi$) is given as follows:
%
\be
\phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}}.
\label{eq:phi}
\ee
%
where the numerator on the right-hand-side denotes the ratio of the fuel (H$_2$)
and oxidizer (O$_2$) at a given condition to the same quantity under stoichiometric
conditions. In this study, computations were performed at fuel-rich conditions,
$\phi$ = 2.0. Homogeneous ignition at constant pressure is simulated using the
TChem software package~\cite{Safta:2011}, and the time required for the rate of
temperature increase to exceed a given threshold, regarded as \emph{ignition delay}
is recorded.

We seek to understand the impact of uncertainty in the
rate-controlling parameters, pre-exponent~($A_i$) and the activation
energy~($E_{a,i}$) on the ignition delay. The $A_i$'s associated with all
reactions and the $E_{a,i}$'s with non-zero nominal estimates
are considered to be uniformly distributed about their nominal estimates provided
in~\cite{Yetter:1991}.  The total number of uncertain inputs is 33 which makes
the present problem computationally challenging due to the large number of 
uncertain parameters.  
To address this challenge, we focus on reducing the dimensionality
of the problem by computing the aforementioned active subspace.
This involves repeated evaluations of the gradient of a model output with
respect to the input parameters. Several numerical techniques are available
for computing the gradient. Perturbation-based techniques
include finite differences and more advanced methods involving
adjoints~\cite{Jameson:1988,Borzi:2011,Alexanderian:2017}. The
adjoint-based method requires a solution of the state equation (forward solve)
and the corresponding adjoint equation. Hence, it is 
limited by the availability of an adjoint solver. Additional model evaluations
at neighboring points are required if finite difference is used which increases
the computational effort. Regression-based techniques aim to estimate the
gradient by approximating the model output using a regression fit. These
are computationally less intensive than the former. However, as expected,
there is a trade-off between computational effort and accuracy in the two
approaches for estimating the gradient.

In this work, we adopt an iterative strategy to reduce the computational effort
associated with active subspace computation. Moreover, we explore two approaches
for estimating the gradient of the ignition delay with respect to the uncertain
rate-controlling parameters: pre-exponent, $A_i$ and the activation energy, $E_{a,i}$,
for $i^{\text{th}}$ reaction in the H$_2$/O$_2$ mechanism. As mentioned above, the
first approach uses finite difference to estimate the gradient and is regarded as
the perturbation-based approach, whereas, the second approach involves a linear
regression-fit to the available set of model evaluations and is regarded as
the regression-based approach.

@@

The main contributions of this paper are as follows: 
\begin{itemize}
\item 
Active subspace
discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving 33
uncertain inputs: The methodology presented in this work
resulted in a 1-dimensional active subspace, indicating immense potential
for computational savings. The presented
analysis can also guide practitioners in other problems of chemical kinetics on using the
method of active subspaces to achieve efficiency in uncertainty propagation.  
\item Comprehensive numerical investigation of the perturbation-based and the regression-based approaches: 
We investigate the suitability of both approaches
for estimating the gradient of ignition delay in the H$_2$/O$_2$ mechanism.
Specifically, we compare resulting
active subspaces, surrogate models, and the ability to approximate global
sensitivity measures through a comprehensive set of numerical experiments. Our
results reveal insight into the merits of the methods as well as
their shortcomings.  
\item Analysis of the links between  
global sensitivity measures: 
By connecting the recent theoretical advances in variance-based and
derivative-based global sensitivity analysis to active subspaces, we provide a complete analysis
of the links between total Sobol' indices, DGSMs, and activity scores for a broad
class of probability distributions. Our analysis is concluded by a result quantifying
approximation errors incurred due to fixing unimportant parameters, deemed so by 
computing their activity scores.

@@

This article is organized as follows. In section~\ref{sub:ac}, a brief
theoretical background on the active subspace methodology is provided. In
section~\ref{sub:gsa}, it is shown that the activity scores provide a
reasonable approximation to the DGSMs especially in a high-dimensional setting.
Additionally, a relationship between the three global sensitivity measures,
namely, the activity scores, DGSMs, and the total Sobol' indices is
established. In section~\ref{sec:method}, a systematic framework for computing
the active subspace is provided. 
Numerical results based on the perturbation-based approach are compared with those
obtained using the regression-based approach.  The active 
subspace is initially computed for a 19-dimensional 
H$_2$/O$_2$ reaction kinetics problem wherein only the $A_i$'s are
considered as uncertain. In this case, the two approaches are observed to yield
consistent results. We further compute the active  for a 33-dimensional H$_2$/O$_2$
reaction kinetics problem in section~\ref{sec:app}. Once again, 
active subspace is found to be 1-dimensional. However, in this case,
it is observed that although the regression-based estimation of the gradient results in a 
reasonable approximation of the mean and mode, the variability in the ignition delay
due to uncertain rate-controlling parameters is under-estimated. Additionally,  
the sensitivity towards one of the most important parameter is not captured.
Hence, a trade-off between efficiency and accuracy between the two approaches is
apparent in the higher-dimensional problem. 
Finally, a summary and discussion based on our findings is included in
section~\ref{sec:conc}.

>>>>

The application problem considered in this work is the
H$_2$/O$_2$ reaction mechanism from~\cite{Yetter:1991}. This mechanism has gained
a lot of attention as a potential source of clean energy for
locomotive applications~\cite{Das:1996}, and more recently in fuel
 cells~\cite{Loges:2008,Cosnier:2016}. 
The mechanism involves 19
reactions including chain reactions, dissociation/recombination reactions, and
formation and consumption of intermediate species; see Table~\ref{tab:kinetics}. 
For each reaction, the reaction rate is assumed to follow an Arrhenius
correlation with temperature:
%
\be
k_i(T) = A_iT^{n_i}\exp(-E_{a,i}/RT), 
\label{eq:rate}
\ee
%
where $A_i$ is the pre-exponent, $n_i$ is the index of $T$, $E_{a,i}$ is the
activation energy corresponding to the $i^{th}$ reaction, and $R$ is the
universal gas constant.  \rebut{The Arrhenius rate law in~\eqref{eq:rate} 
is often interpreted in a logarithmic form as follows:
%
\be
\log(k_i) = \log(A_i) + n_i\log(T) - E_{a,i}/RT. 
\label{eq:ratelog}
\ee
}
%
%
\begin{table}[htbp]
\renewcommand*{\arraystretch}{0.9}
\begin{center}
\begin{tabular}{llll}
\toprule
Reaction \#     & Reaction &&\\
\bottomrule
$\mathcal{R}_1$ & H + O$_2$          & $\rightleftharpoons$ & O + OH \\
$\mathcal{R}_2$ & O + H$_2$          & $\rightleftharpoons$ & H + OH \\
$\mathcal{R}_3$ & H$_2$ + OH         & $\rightleftharpoons$ & H$_2$O + H \\
$\mathcal{R}_4$ & OH + OH            & $\rightleftharpoons$ & O + H$_2$O \\
$\mathcal{R}_5$ & H$_2$ + M          & $\rightleftharpoons$ & H + H + M \\
$\mathcal{R}_6$ & O + O + M          & $\rightleftharpoons$ & O$_2$ + M \\
$\mathcal{R}_7$ & O + H + M          & $\rightleftharpoons$ & OH + M \\
$\mathcal{R}_8$ & H + OH +M          & $\rightleftharpoons$ & H$_2$O + M \\
$\mathcal{R}_9$ & H + O$_2$ + M      & $\rightleftharpoons$ & HO$_2$ + M \\
$\mathcal{R}_{10}$ & HO$_2$ + H      & $\rightleftharpoons$ & H$_2$ + O$_2$ \\
$\mathcal{R}_{11}$ & HO$_2$ + H      & $\rightleftharpoons$ & OH + OH \\
$\mathcal{R}_{12}$ & HO$_2$ + O      & $\rightleftharpoons$ & O$_2$ + OH \\
$\mathcal{R}_{13}$ & HO$_2$ + OH     & $\rightleftharpoons$ & H$_2$O + O$_2$ \\
$\mathcal{R}_{14}$ & HO$_2$ + HO$_2$ & $\rightleftharpoons$ & H$_2$O$_2$ + O$_2$ \\
$\mathcal{R}_{15}$ & H$_2$O$_2$ + M  & $\rightleftharpoons$ & OH + OH + M \\
$\mathcal{R}_{16}$ & H$_2$O$_2$ + H  & $\rightleftharpoons$ & H$_2$O + OH \\
$\mathcal{R}_{17}$ & H$_2$O$_2$ + H  & $\rightleftharpoons$ & HO$_2$ + H$_2$ \\
$\mathcal{R}_{18}$ & H$_2$O$_2$ + O  & $\rightleftharpoons$ & OH + HO$_2$ \\
$\mathcal{R}_{19}$ & H$_2$O$_2$ + OH & $\rightleftharpoons$ & HO$_2$ + H$_2$O \\
\bottomrule
\end{tabular}
\end{center}
\caption{Reaction mechanism for H$_2$/O$_2$ from~\cite{Yetter:1991}}.
\label{tab:kinetics}
\end{table}
%
The global reaction associated with the H$_2$/O$_2$ mechanism can
be considered as follows:
\be
2\text{H}_2 + \text{O}_2 \rightarrow 2\text{H}_2\text{O}.
\label{eq:global}
\ee 
The equivalence ratio ($\Phi$) is given as follows:
%
\be
\Phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}}.
\label{eq:phi}
\ee
%
where the numerator on the right-hand-side denotes the ratio of the fuel (H$_2$)
and oxidizer (O$_2$) at a given condition to the same quantity under stoichiometric
conditions. In this study, computations were performed at fuel-rich conditions,
$\Phi$ = 2.0. Homogeneous ignition at constant pressure is simulated using the
TChem software package~\cite{Safta:2011} \rebut{using an initial pressure, $P_0$ = 1 atm and
initial temperature, $T_0$ = 900~K.} The time required for the rate of
temperature increase to exceed a given threshold, regarded as \emph{ignition delay}
is recorded.

We seek to understand the impact of uncertainty in the
rate-controlling parameters, pre-exponent~($A_i$) and the activation
energy~($E_{a,i}$) \rebut{as well as the initial pressure, temperature, and the
equivalence ratio} on the ignition delay. The $A_i$'s associated with all
reactions and the $E_{a,i}$'s with non-zero nominal estimates
are considered to be uniformly distributed about their nominal estimates provided
in~\cite{Yetter:1991}. \rebut{The initial conditions are also considered to be uniformly
distributed about their respective aforementioned values.} 
The total number of uncertain inputs is \rebut{36} which makes
the present problem computationally challenging due to the large number of 
uncertain parameters \rebut{in addition to the initial conditions}.  
To address this challenge, we focus on reducing the dimensionality
of the problem by computing the active subspace.
This involves repeated evaluations of the gradient of a model output with
respect to the input parameters. Several numerical techniques are available
for computing the gradient. Perturbation-based techniques
include finite differences and more advanced methods involving
adjoints~\cite{Jameson:1988,Borzi:2011,Alexanderian:2017}. The
adjoint-based method requires a solution of the state equation (forward solve)
and the corresponding adjoint equation. Hence, it is 
limited by the availability of an adjoint solver. Additional model evaluations
at neighboring points are required if finite difference is used which increases
the computational effort. Regression-based techniques on the other hand,
aim to estimate the
gradient by approximating the model output using a regression fit. These
are computationally less intensive than the former. However, as expected,
there is a trade-off between computational effort and accuracy in the two
approaches for estimating the gradient.

In this work, we adopt an iterative strategy to reduce the computational effort
associated with active subspace computation. Moreover, we explore two approaches
for estimating the gradient of the ignition delay with respect to the uncertain
rate-controlling parameters: pre-exponent, $A_i$ and the activation energy, $E_{a,i}$,
for $i^{\text{th}}$ reaction in the H$_2$/O$_2$ mechanism \rebut{as well as
the initial conditions: $P_0$, $T_0$, and $\Phi_0$. Note that the equivalence ratio
corresponding to the initial molar ratios of $H_2$ and O$_2$ is denotes as $\Phi_0$.}
As mentioned above, the
first approach uses finite difference to estimate the gradient and is regarded as
the perturbation approach, whereas, the second approach involves a linear
regression-fit to the available set of model evaluations and is regarded as
the regression approach.

@@

The main contributions of this paper are as follows: 
\begin{itemize}
\item 
Active subspace
discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving \rebut{36}
uncertain inputs: The methodology presented in this work
\rebut{successfully demonstrated that a 1-dimensional active subspace can reasonably 
approximate the uncertainty in the QoI}, indicating immense potential
for computational savings. The presented
analysis can also guide practitioners in other problems of chemical kinetics on using the
method of active subspaces to achieve efficiency in uncertainty propagation.  
\item Comprehensive numerical investigation of the perturbation and the regression approaches: 
We investigate the suitability of both approaches
for estimating the gradient of ignition delay in the H$_2$/O$_2$ mechanism.
Specifically, we compare resulting
active subspaces, surrogate models, and the ability to approximate global
sensitivity measures through a comprehensive set of numerical experiments. Our
results reveal insight into the merits of the methods as well as
their shortcomings.  
\item Analysis of the links between  
global sensitivity measures: 
By connecting the recent theoretical advances in variance-based and
derivative-based global sensitivity analysis to active subspaces, we provide a complete analysis
of the links between total Sobol' indices, DGSMs, and activity scores for a broad
class of probability distributions. Our analysis is concluded by a result quantifying
approximation errors incurred due to fixing unimportant parameters, deemed so by 
computing their activity scores.

@@

This article is organized as follows. In section~\ref{sub:ac}, a brief
theoretical background on the active subspace methodology is provided. In
section~\ref{sub:gsa}, it is shown that the activity scores provide a
reasonable approximation to the DGSMs especially in a high-dimensional setting.
Additionally, a relationship between the three global sensitivity measures,
namely, the activity scores, DGSMs, and the total Sobol' indices is
established. In section~\ref{sec:method}, a systematic framework for computing
the active subspace is provided. 
Numerical results based on the perturbation approach are compared with those
obtained using the regression approach.  The active 
subspace is initially computed for a 19-dimensional 
H$_2$/O$_2$ reaction kinetics problem wherein only the $A_i$'s are
considered as uncertain. We further compute the active subspace
for a 36-dimensional H$_2$/O$_2$ reaction kinetics problem in section~\ref{sec:app}.
\rebut{For both settings, the convergence characteristics and the predictive accuracy
of the two approaches is compared for a given amount of computational effort. The two
approaches are observed to yield consistent results, and a 1-dimensional active subspace
is observed to capture the uncertainty in the ignition delay.}
Finally, a summary and discussion based on our findings is included in
section~\ref{sec:conc}.
